---
title: "Time sequences"
author: "Pep Porrà"
date: "March, 21st 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

To give examples of time sequence management using the package {xts}.

## Method

We will define several time sequences and operate with them to give useful tricks.

## Dates sequences definition

Use `seq.Date(from, to, by, length.out, along.with)` to define sequences of dates

The `by` parameter can be "day", "week", "month", "quarter" or "year" with a positive or negative integer preceding it and a space. Also a number that is taken as days.

We define first a sequence of 5 dates starting on 2000-01-01 every 10 days

```{r}
tseq1 <- seq(from = as.Date('2000-01-01'), by = 10, len = 5)
tseq1
```

Same as tseq1 but every 2 weeks until 2000-02-28

```{r}
(tseq2 <- seq.Date(as.Date("2000-01-01"), as.Date("2000-02-28"), by = "2 weeks"))
```

Third sequence is the first days of the first 5 years of 21st century

```{r}
tseq3 <- seq.Date(from = as.Date('2000-01-01'), by = 'years', len = 5)
tseq3
```

Last day of every quarter, in 2002

```{r}
tseq4 <- seq.Date(from = as.Date('2002-03-31'),
  to = as.Date("2002-12-31"), by = "quarter")
tseq4
```

Last days of the first 5 month of 2004
```{r}
tseq5 <- seq.Date(from = as.Date('2004-01-31'),
  by = "1 months", len = 5)
tseq5 
tseq5 <- seq.Date(from = as.Date('2004-02-01'),
  by = "1 months", len = 5) - 1
tseq5
```

Last day of February of for years 2002 to 2007

```{r}
tseq6 <- seq.Date(from = as.Date('2001-02-28'), to = as.Date("2007-03-01"),
  by = "1 year")
tseq6
tseq6 <- seq.Date(from = as.Date('2001-03-01'), to = as.Date("2007-03-01"),
  by = "1 year") - 1
tseq6
```

